
```{r dependencies}
library(tidyverse)
library(readxl)
library(janitor)
library(skimr)
```

## Read
```{r load}
august_enrollments <-
  read_excel("./data/august-enrolments-by-school.xlsx",
             sheet = 3,
             skip = 13)

august_enrollments
```

The column names have annoying spaces. I'm just looking at secondary schools.

## Subset
```{r subset}
august_enrollments_hs <-
  august_enrollments %>%
  janitor::clean_names() %>%
  filter(school_category == "Secondary",
         school_type != "Special School")

```

## Munge

```{r munge}
august_cohorts_hs <-
  august_enrollments_hs %>%
  extract("year_level",
          into = "grade",
          regex = "([0-9]+)",
          convert = TRUE,
          remove = FALSE
          ) %>%
  mutate(cohort = collection_year - grade)

hs_cohorts_aug <-
  august_cohorts_hs %>%
  group_by(cohort) %>%
  filter(min(collection_year) <=  2014)

skim(hs_cohorts_aug)
```
There's some grades that are NA. Need to go back to `year_level`.

```{r}
august_enrollments_hs %>%
  pull(year_level) %>%
  unique()


hs_cohorts_aug <-
  august_cohorts_hs %>%
  filter(year_level != "Special") %>%
  group_by(cohort) %>%
  filter(min(collection_year) <=  2014,
         max(grade) == 12,
         min(grade) == 8) %>%
  ungroup()

skim(hs_cohorts_aug)
 ```

Got rid of the NA.

## Cohort analysis
We have one complete cohort "2005" starting in year 8 in 2013 and finishing yr
12 in 2017. A quick way to 'normalise' each cohort is to look at the enrolments
for each year minus the enrolments in yr 8.

### Normalise

```{r}
starting_enrolments <-
  hs_cohorts_aug %>%
  filter(grade == 8) %>%
  rename(initial_enrolments = ft_enrolments) %>%
  select(school_name, cohort, initial_enrolments)

hs_cohorts_aug %>%
  left_join(starting_enrolments) %>%
  mutate(enrolments_diff = ft_enrolments - initial_enrolments) %>%
  ggplot(mapping = aes(x = grade, y = enrolments_diff)) +
  geom_line(aes(group = school_name,
                colour = region))
```

