
```{r dependencies}
library(tidyverse)
library(readxl)
library(janitor)
library(skimr)
library(leaflet)
library(scico)
```

## Read
```{r load}
august_enrollments <-
  read_excel("./data/august-enrolments-by-school.xlsx",
             sheet = 3,
             skip = 13)

august_enrollments
```

The column names have annoying spaces. I'm just looking at secondary schools.

## Subset
```{r subset}
august_enrollments_hs <-
  august_enrollments %>%
  janitor::clean_names() %>%
  filter(school_category == "Secondary",
         school_type != "Special School")

```

## Munge

```{r munge}
august_cohorts_hs <-
  august_enrollments_hs %>%
  extract("year_level",
          into = "grade",
          regex = "([0-9]+)",
          convert = TRUE,
          remove = FALSE
          ) %>%
  mutate(cohort = collection_year - grade)

hs_cohorts_aug <-
  august_cohorts_hs %>%
  group_by(cohort) %>%
  filter(min(collection_year) <=  2014)

skim(hs_cohorts_aug)
```
There's some grades that are NA. Need to go back to `year_level`.

```{r}
august_enrollments_hs %>%
  pull(year_level) %>%
  unique()


hs_cohorts_aug <-
  august_cohorts_hs %>%
  filter(year_level != "Special") %>%
  group_by(cohort, school_name) %>%
  filter(min(collection_year) <=  2014,
         max(grade) == 12,
         min(grade) == 8) %>%
  ungroup()

skim(hs_cohorts_aug)
 ```

Got rid of the NA.

## Cohort analysis
We have one complete cohort "2005" starting in year 8 in 2013 and finishing yr
12 in 2017. A quick way to 'normalise' each cohort is to look at the enrolments
for each year minus the enrolments in yr 8.

### Normalise

```{r}
starting_enrolments <-
  hs_cohorts_aug %>%
  filter(grade == 8) %>%
  rename(initial_enrolments = ft_enrolments) %>%
  select(school_name, cohort, initial_enrolments)

hs_cohorts_aug <-
  hs_cohorts_aug %>%
  left_join(starting_enrolments) %>%
  mutate(enrolments_diff = ft_enrolments / initial_enrolments)

hs_cohorts_aug %>%
  ggplot(mapping = aes(x = grade, y = enrolments_diff)) +
  geom_line(aes(group = school_name,
                colour = region))
```

## Geocoding

Using `opencage`: https://github.com/ropensci/opencage

```{r}
library(opencage)

## Places
schools <- paste(unique(hs_cohorts_aug$school_name), "QLD")

geocode_results <-
  map(schools, ~opencage_forward(placename = ., countrycode = "AU"))

write_rds(geocode_results, "./data/geocode_results.Rds")
```
## Merge Geo Data From QLD gov open data

```{r}
qld_schools <-
  read_excel("./data/state-and-non-state-school-listing.XLSX",
             sheet = 2,
             skip = 10,
             guess_max = 1500
             ) %
         n_names() %>%
  select(school_name,
         school_code,
        federal_electorate,
         state_electorate,
         latitude,
         longitude)
## check for non matches

hs_cohorts_aug %>%
  anti_join(qld_schools)

hs_cohorts_aug <-
  hs_cohorts_aug %>%
  left_join(qld_schools)

```

That worked a bit better.

## Simple spatial vis

A leaflet map with circles representing schools coloured by proportion in of initial enrolments present in yr 12 cohort.

```{r}
hs_cohorts_aug_12prop <- 
  hs_cohorts_aug %>%
  filter(grade == 12) %>%
  mutate(prop_12 = ft_enrolments/initial_enrolments) %>%
  rename(lat = latitude, lng = longitude) %>%
  mutate(lat = as.numeric(lat), lng = as.numeric(lng))

circle_palette <-
  colorNumeric(scico(500, palette = "vik"),
               hs_cohorts_aug_12prop$prop_12,
               reverse = TRUE)


leaflet(data = hs_cohorts_aug_12prop) %>%
  addTiles() %>%
  addCircleMarkers(color = ~circle_palette(prop_12),
                   stroke = FALSE,
                   fillOpacity = 0.6
                   ) %>%
  addLegend("bottomright",
            pal = circle_palette,
            values = ~prop_12
          )
```
copy-pasta for yr 10 cohort.
```{r}
hs_cohorts_aug_10prop <- 
  hs_cohorts_aug %>%
  filter(grade == 10) %>%
  mutate(prop_10 = ft_enrolments/initial_enrolments) %>%
  rename(lat = latitude, lng = longitude) %>%
  mutate(lat = as.numeric(lat), lng = as.numeric(lng))

circle_palette <-
  colorNumeric(scico(500, palette = "vik"),
               hs_cohorts_aug_10prop$prop_10,
               reverse = TRUE)


leaflet(data = hs_cohorts_aug_10prop) %>%
  addTiles() %>%
  addCircleMarkers(color = ~circle_palette(prop_10),
                   stroke = FALSE,
                   fillOpacity = 0.6
                   ) %>%
  addLegend("bottomright",
            pal = circle_palette,
            values = ~prop_10
          )
```
